---
title: "Working With Data"
format: 
 pdf:
  number-sections: true
  toc: true
  pdf-engine: xelatex
editor: visual
---

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

##Library

```{r}
library(tidyverse)
library(descr)
library(knitr)
library(dplyr)
library(haven)
library(ggplot2)
library(Hmisc)
library(readr)
library(car)
```

##Data Set Load

```{r}
file_path <- file.path("..", "data", "36168-0001-Data.sav")
df <- read_sav(file_path)
```

##DATA TABLE

```{r}
head(df)
```

##FREQ

```{r}
freq(as.ordered(df$sex), plot = FALSE)
```

This table shows that more female participants went through abuse more than that of the men amount of men whpo participated.

```{r}
freq(as.ordered(df$mooddx), plot = FALSE)
```

This table shows the amount of people with mood disorders and there are more people with than without.

```{r}
freq(as.ordered(df$race), plot = FALSE)
```

This table shows the amount of white, black and other races that went through abuse. White participants were shown to be the ones to go through abuse.

##Data Management

```{r}
#df$race[df$race == 1 | df$race == 2 | df$race == 3 | df$race == 5 | df$race == 6] <- 1
#df$race[df$race == 4 | df$race == 7 | df$race == 8 | df$race == 9]
```

```{r}
df$RACE[is.na(df$RACE)] <- 0
```

```{r}
df <- janitor::clean_names(df)
```

```{r}
names (df)
```

```{r}
#df <- df %>%
  #age_group = case_when()
      #age < 30 ~ "Under 30",
      #age >= 30 & age < 60 ~ "30-59",
      #age >= 60 ~ "60+"
    #sex = recode(as.character(sex
                 "1" = "Male"
                 "2" = "Female"
```

```{r}
#df_summary <- df %>%
  #group_by(sex, age_group) %>%
  #summarise(
    #mean_depression = mean(scl_dep, na.rm = TRUE),
    #count = n()
```

```{r}
#df_summary <- df %>%
  #group_by(sex, age_group) %>%
  #summarise(
    #mean_depression = mean(scl_dep, na.rm = TRUE),
    #count = n(),
    #.groups = "drop"
```

```{r plot_depression_by_age_sex}
#ggplot(df_summary, aes(x = age_group, y = mean_depression, fill = sex)) +
  #geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Mean Depression Scores by Age Group and Sex",
    x = "Age Group",
    y = "Mean Depression Score"
  ) +
  theme_minimal()
```

##Univariate
First Graph representing race

```{r}
ggplot(df, aes(x = race)) + geom_bar()
```

```{r}
ggplot(df, aes(x = age, y = scl_gsi)) +
  geom_point(alpha = 0.6) +
  labs(title = "Age vs. Global Severity Index (SCL-GSI)", x = "Age", y = "SCL-GSI") +
  theme_minimal()
```

Second graph representing age

```{r}
ggplot(df, aes(x = age)) + geom_bar()
```

```{r}
ggplot(df, aes(x = age, y = scl_gsi)) +
  geom_point(alpha = 0.6) +
  labs(title = "Age vs. Global Severity Index (SCL-GSI)", x = "Age", y = "SCL-GSI") +
  theme_minimal()
```

## Chi-Squared

```{r}
table(df$sex)
table(df$dx)
```

```{r}
table(df$sex, df$dx)
```

```{r}
chisq.test(table(df$sex, df$dx))
```

```{r}
table(df$sex, df$abuse)
chisq.test(table(df$sex, df$abuse))

```

```{r}
df <- df %>%
  mutate(abuse = as.factor(abuse))

```

##T-Test

```{r}
t.test(age ~ sex, data = df)
```

```{r}
df <- df %>%
  mutate(
    sex = as.factor(sex),
    abuse = as.factor(abuse)
  )
```

```{r}
t.test(df$age, mu = 30)

```

```{r}
ggplot(df, aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "#69b3a2", alpha = 0.7, color = "white") +
  geom_vline(xintercept = 30, linetype = "dashed", color = "red", size = 1.2) +
  labs(
    title = "Age Distribution",
    x = "Age",
    y = "Frequency"
  ) +
  theme_minimal()
```

People affected by abuse mostly have effects ages 30 and 40.

##ANOVA

```{r}
set.seed(123)
```

```{r}
aov_age_sex <- aov(age ~ sex, data = df)
```

```{r}
summary(aov_age_sex)
```

This shows that there is no significant correlation between sex and age of people affected by abuse.

##Correlation

```{r}
cor.test(df$age, df$scl_gsi, use = "complete.obs")
```

```{r}
ggplot(data = df, aes(x = age, y = scl_gsi)) + 
  geom_point(color = "purple") +
  theme_bw() +
  labs(x = "Participant Age (years)", y = "Global Severity Index (scl_gsi)") +
  stat_smooth(method = lm)

```

There's no real link between age and psychological distress
